TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-05-05 paper/draft.tex
add comments

A number of M dwarf studies have already been conducted in \apogee, particularly towards measuring reliable fundamental atmospheric parameters and kinematic measurements. \citealt{Schmidt:2016} -- analyzed trends and reliablility of ASPCAP parameters for K/M dwarfs. \citealt{Souto:2017} and \citealt{Souto:2018} -- modeled three exoplanet hosting M dwarfs (Kepler-138, Kepler-168 and Ross-128), determining Teff/logg/metallicity + 13 elemental abundances. \citealt{Desphande:2013} -- modeled rotational and radial velocities for 200+ M dwarfs, later extended by \citealt{Gilhool:2018} to 700+ M dwarfs. \citealt{Rajpurohit:2018} -- tested BTSettl model grids on 45 M dwarfs, estimated Teff/logg/metallicity). \citealt{Skinner:2018} - identified and measured mass ratios and radial velocities for 44 M dwarf binaries.

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
This percent-level accuracy of the model in the data domain is also demonstrated...
% \color{gcolor}{HOGG: Adam says he doesn't think this is right: if $\chi^2/n>>1$ then the model is NOT an accurate representation of the data.}\color{black} \\
\color{gcolor}{BIRKY: Can you demonstrate this by showing the RMS residuals, putting it in as a figure, and putting this paragraph back to me? I am imagining a figure that shows a scatter plot of T vs RMS residual, colored by metallicity. Or something like that.}\color{black}

>>>>

\color{gcolor}{JB: I've moved this paragraph from the intro to here. Will expand this... A number of M dwarf studies have already been conducted in \apogee, particularly towards measuring reliable fundamental atmospheric parameters and kinematic measurements. \citealt{Schmidt:2016} -- analyzed trends and reliablility of ASPCAP parameters for K/M dwarfs. \citealt{Souto:2017} and \citealt{Souto:2018} -- modeled three exoplanet hosting M dwarfs (Kepler-138, Kepler-168 and Ross-128), determining Teff/logg/metallicity + 13 elemental abundances. \citealt{Desphande:2013} -- modeled rotational and radial velocities for 200+ M dwarfs, later extended by \citealt{Gilhool:2018} to 700+ M dwarfs. \citealt{Rajpurohit:2018} -- tested BTSettl model grids on 45 M dwarfs, estimated Teff/logg/metallicity). \citealt{Skinner:2018} - identified and measured mass ratios and radial velocities for 44 M dwarf binaries.}\color{black}

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude.
They are much higher.
This discrepancy flows from the fact that---while the spectral model is good at the percent
level---the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.
This percent-level accuracy of the model in the data domain is also demonstrated... \\

\color{gcolor}{BIRKY: Can you demonstrate this by showing the RMS residuals, putting it in as a figure, and putting this paragraph back to me? I am imagining a figure that shows a scatter plot of T vs RMS residual, colored by metallicity. Or something like that.}\color{black} \\

\color{gcolor}{HOGG: Adam's comment: if $\chi^2/n>>1$ then the model is NOT an accurate representation of the data.}\color{black} \\
